A Hybrid Neutronics Method for Time-Dependent Control Rod Modeling in Molten Salt Reactors

### 2024-06-20 moltres.tex
Fix typos and add numerical implementation section

\section{Description of Moltres}

@@

Moltres draws many advantages from being developed on MOOSE
\cite{permann_moose_2020}. MOOSE is an open-source, finite-element,
multiphysics framework developed at \gls{INL}. The framework provides a
user-friendly interface for developing multiphysics software through
\gls{OOP} in \texttt{C++} to modularize various
functions relevant to finite-element multiphysics solvers. In this approach,
MOOSE and MOOSE-based applications break down \glspl{PDE} into individual terms
and store them as individual \texttt{C++ objects} referred to as
\texttt{Kernels}. These \texttt{Kernels} contain functions for calculating
their weak form residual and Jacobian
contributions and other relevant functions required to solve a given
\gls{PDE}. \gls{OOP} in MOOSE simplifies software development
since developers can write new \texttt{Kernels} as child classes in
\texttt{C++} derived from existing \texttt{Kernels} (base classes), which share
similar physics to inherit common functions.
The same philosophy applies to all other systems in MOOSE, such as
the \texttt{BCs}, \texttt{Materials}, and \texttt{Postprocessor}
systems for handling relevant boundary conditions, material properties, and
postprocessing calculations, respectively. Overall, this approach also saves
researchers time and effort since they are unencumbered by the technical details
and complexities of programming efficient computational tools for
numerical analysis.

@@

Moltres solves the multigroup neutron diffusion equations for the neutron
flux solution within the problem domain. These equations are derived from the
neutron transport equation in the diffusion-dominated limit with Fick's law of
diffusion. They are further simplified by discretizing the continuous neutron energy
variable into a finite number of energy groups \cite{bell_nuclear_1970,
duderstadt_nuclear_1976}. The time-dependent multigroup neutron
diffusion equations with $G$ energy groups and $I$ \gls{DNP}
groups are given by:
%
\begin{align}
    \frac{1}{v_g} \frac{\partial \phi_g}{\partial t} =& \nabla \cdot D_g
    \nabla \phi_g - \Sigma^r_g \phi_g +
    \sum^G_{g' \neq g} \Sigma^s_{g' \rightarrow g} \phi_{g'} \nonumber \\
    &+ \chi^p_g \sum^G_{g'=1} \left( 1-\beta \right) \nu \Sigma^f_{g'}
    \phi_{g'} + \chi^d_g \sum^I_i \lambda_i C_i \label{eq:neutron} %\\
%    %
%    \shortintertext{where}
%    v_g =& \text{ average speed of neutrons in group $g$,} 
%    \nonumber \\
%    \phi_g =& \text{ neutron flux in group $g$,}
%    \nonumber \\
%    t =& \text{ time,} \nonumber \\
%    D_g =& \text{ diffusion coefficient of neutrons in group $g$,} \nonumber \\
%    \Sigma^r_g =& \text{ macroscopic removal cross section for} \nonumber \\
%    &\text{ neutrons from group $g$,} \nonumber \\
%    \Sigma^s_{g' \rightarrow g} =& \text{ macroscopic scattering cross section
%    for neutrons from} \nonumber \\
%    &\text{ groups $g'$ to $g$,} \nonumber \\
%    \chi^p_g =& \text{ prompt fission spectrum for neutrons in group $g$,} \nonumber \\
%    G =& \text{ total number of discrete neutron groups,} \nonumber \\
%    \nu_g =& \text{ average number of neutrons produced per fission,} \nonumber
%    \\
%    \Sigma^f_{g} =& \text{ macroscopic fission cross section for neutron}
%    \nonumber \\
%    &\text{ in group $g$,} \nonumber \\
%    \chi^d_g =& \text{ delayed fission spectrum for neutrons in group $g$,} \nonumber \\
%    I =& \text{ total number of \gls{DNP} groups,} \nonumber \\
%    \beta =& \text{ total delayed neutron fraction.} \nonumber
\end{align}

Despite forming only around 0.7\% of all neutrons emitted, delayed neutrons
play outsized roles in reactor kinetics. The relatively long half-lives of
\glspl{DNP} give reactor operators ample time in adequately designed reactors
to control reactor power output and intervene in case of power excursions.
The precursor concentration balance equations for $I$ precursor
groups are given by:
%
\begin{align}
    \frac{\partial C_i}{\partial t} =& \beta_i \sum^G_{g'=1} \nu \Sigma^f_{g'}
    \phi_{g'} - \lambda_i C_i - \vec{u} \cdot \nabla C_i + \nabla \cdot
    D_{\text{P}} \nabla C_i \label{eq:precursor} %\\
    %
%    \shortintertext{where}
%    \beta_i =& \text{ delayed neutron fraction of precursor group $i$,}
%    \nonumber \\
%    \lambda_i =& \text{ average decay constant of delayed neutron} \nonumber \\
%    &\text{ precursors in precursor group $i$,} \nonumber \\
%    C_i =& \text{ concentration of \gls{DNP}s in}
%    \nonumber \\
%    &\text{ precursor group $i$,} \nonumber \\
%    \vec{u} =& \text{ molten salt flow velocity vector,}
%    \nonumber \\
%    D_{\text{P}} =& \text{ effective diffusion coefficient of the delayed}
%    \nonumber \\
%    &\text{ neutron precursors.} \nonumber
\end{align}

@@

\begin{align}
    \text{Momentum equation: } \rho \frac{\partial \vec{u}}{\partial t} =&
    -\rho (\vec{u}
    \cdot \nabla) \vec{u} - \nabla p + \mu \nabla^2 \vec{u}
    + \rho \alpha \vec{g} \left(T - T_{\text{ref}} \right)
    \label{eq:momemtum}
    \shortintertext{and}
    \text{Mass equation: } \nabla \cdot \vec{u} =& 0
    \label{eq:divergence}
%    \shortintertext{where}
%    \rho =& \text{ fluid density,} \nonumber \\
%    p =& \text{ pressure,} \nonumber \\
%    \mu =& \text{ dynamic viscosity,} \nonumber \\
%    \alpha =& \text{ coefficient of thermal expansion,} \nonumber \\
%    \vec{g} =& \text{ gravitational force vector,} \nonumber
%    \\
%    T =& \text{ fluid temperature,} \nonumber \\
%    T_{\text{ref}} =& \text{ reference temperature at which the nominal}
%    \nonumber \\
%    &\text{ density is provided.} \nonumber
%    \nonumber
\end{align}

@@

\begin{align}
    \rho c_{p} \frac{\partial T}{\partial t} =& - \rho c_p \vec{u}
    \cdot \nabla T + \nabla \cdot \left(k \nabla T \right) + Q_f - Q_s
    \label{eq:temp}
    \shortintertext{and}
    Q_f =& \sum^G_{g=1} \epsilon_g \Sigma_g^f \phi_g \label{eq:heat-source}
    \shortintertext{where}
%    c_p =& \text{ specific heat capacity of molten salt,} \nonumber \\
%    k =& \text{ effective thermal conductivity of molten salt,} \nonumber \\
%    Q_f =& \text{ fission heat source,} \nonumber \\
%    \epsilon_g =& \text{ average fission energy released by neutrons in group
%    $g$,} \nonumber \\
    Q_s =& \text{ heat sink/removal.} \nonumber
\end{align}

@@

Moltres also accounts for the decay of
\glspl{DNP} outside the active core region by simulating its flow in a
separate 1-D pipe geometry. This external loop pipe calculation is tightly
coupled to the active core simulation through Picard iterations in MOOSE's
MultiApp functionality and inlet/outlet boundary values.
The external loop region is assumed to be subcritical to minimize neutron
irradiation upon heat exchangers, pumps, and other equipment. Therefore, the
only significant neutronic-related phenomena are the drift and decay of
\glspl{DNP}. The governing equation for the \glspl{DNP} is:
%
\begin{align}
    \frac{\partial C_i}{\partial t} =& - \lambda_i C_i - u
    \frac{\partial C_i}{\partial x}.
    \label{eq:dnploop}
\end{align}
%
Equation \ref{eq:dnploop} is derived from equation \ref{eq:precursor} by
removing the fission \gls{DNP} source and diffusion terms and reducing the
dimensionality from 3-D to 1-D.

>>>>

\section{Description of Moltres} \label{sec:moltres-description}

@@

Moltres draws many advantages from being developed on MOOSE
\cite{giudicelli_30_2024}. MOOSE is an open-source, finite-element,
multiphysics framework developed at \gls{INL}. The framework provides a
user-friendly interface for developing multiphysics software through
\gls{OOP} in \texttt{C++} to modularize various
functions relevant to finite-element multiphysics solvers. In this approach,
MOOSE and MOOSE-based applications break down \glspl{PDE} into individual terms
and store them as individual \texttt{C++ objects} referred to as
\texttt{Kernels}. These \texttt{Kernels} contain functions for calculating
their weak form residual and Jacobian
contributions and other relevant functions required to solve a given
\gls{PDE}. \gls{OOP} in MOOSE simplifies software development
since developers can write new \texttt{Kernels} as child classes in
\texttt{C++} derived from existing \texttt{Kernels} (base classes), which share
similar physics to inherit common functions.
The same philosophy applies to all other systems in MOOSE, such as
the \texttt{BCs}, \texttt{Materials}, and \texttt{Postprocessor}
systems for handling relevant boundary conditions, material properties, and
postprocessing calculations, respectively. Overall, this approach also saves
researchers time and effort since they are unencumbered by the technical details
and complexities of programming efficient computational tools for
numerical analysis.

@@

Moltres solves the multigroup neutron diffusion equations for the neutron
flux solution within the problem domain. These equations are derived from the
neutron transport equation in the diffusion-dominated limit with Fick's law of
diffusion. They are further simplified by discretizing the continuous neutron energy
variable into a finite number of energy groups \cite{bell_nuclear_1970,
duderstadt_nuclear_1976}. The time-dependent multigroup neutron
diffusion equations with $G$ energy groups and $I$ \gls{DNP}
groups are given by:
%
\begin{align}
    \frac{1}{v_g} \frac{\partial \phi_g}{\partial t} =& \nabla \cdot D_g
    \nabla \phi_g - \Sigma^r_g \phi_g +
    \sum^G_{g' \neq g} \Sigma^s_{g' \rightarrow g} \phi_{g'} \nonumber \\
    &+ \chi^p_g \sum^G_{g'=1} \left( 1-\beta \right) \nu \Sigma^f_{g'}
    \phi_{g'} + \chi^d_g \sum^I_i \lambda_i C_i \label{eq:neutron} %\\
    %
    \shortintertext{where}
    v_g =& \text{ average speed of neutrons in group $g$,} 
    \nonumber \\
    \phi_g =& \text{ neutron flux in group $g$,}
    \nonumber \\
    t =& \text{ time,} \nonumber \\
    D_g =& \text{ diffusion coefficient of neutrons in group $g$,} \nonumber \\
    \Sigma^r_g =& \text{ macroscopic removal cross section for} \nonumber \\
    &\text{ neutrons from group $g$,} \nonumber \\
    \Sigma^s_{g' \rightarrow g} =& \text{ macroscopic scattering cross section
    for neutrons from} \nonumber \\
    &\text{ groups $g'$ to $g$,} \nonumber \\
    \chi^p_g =& \text{ prompt fission spectrum for neutrons in group $g$,} \nonumber \\
    G =& \text{ total number of discrete neutron groups,} \nonumber \\
    \nu_g =& \text{ average number of neutrons produced per fission,} \nonumber
    \\
    \Sigma^f_{g} =& \text{ macroscopic fission cross section for neutron}
    \nonumber \\
    &\text{ in group $g$,} \nonumber \\
    \chi^d_g =& \text{ delayed fission spectrum for neutrons in group $g$,} \nonumber \\
    I =& \text{ total number of \gls{DNP} groups,} \nonumber \\
    \beta =& \text{ total delayed neutron fraction.} \nonumber
\end{align}

Despite forming only around 0.7\% of all neutrons emitted, delayed neutrons
play outsized roles in reactor kinetics. The relatively long half-lives of
\glspl{DNP} give reactor operators ample time in adequately designed reactors
to control reactor power output and intervene in case of power excursions.
The precursor concentration balance equations for $I$ precursor
groups are given by:
%
\begin{align}
    \frac{\partial C_i}{\partial t} =& \beta_i \sum^G_{g'=1} \nu \Sigma^f_{g'}
    \phi_{g'} - \lambda_i C_i - \vec{u} \cdot \nabla C_i + \nabla \cdot
    D_{\text{P}} \nabla C_i \label{eq:precursor} %\\
    %
    \shortintertext{where}
    \beta_i =& \text{ delayed neutron fraction of precursor group $i$,}
    \nonumber \\
    \lambda_i =& \text{ average decay constant of delayed neutron} \nonumber \\
    &\text{ precursors in precursor group $i$,} \nonumber \\
    C_i =& \text{ concentration of \gls{DNP}s in}
    \nonumber \\
    &\text{ precursor group $i$,} \nonumber \\
    \vec{u} =& \text{ molten salt flow velocity vector,}
    \nonumber \\
    D_{\text{P}} =& \text{ effective diffusion coefficient of the delayed}
    \nonumber \\
    &\text{ neutron precursors.} \nonumber
\end{align}

@@

\begin{align}
    \text{Momentum equation: } \rho \frac{\partial \vec{u}}{\partial t} =&
    -\rho (\vec{u}
    \cdot \nabla) \vec{u} - \nabla p + \mu \nabla^2 \vec{u}
    + \rho \alpha \vec{g} \left(T - T_{\text{ref}} \right)
    \label{eq:momemtum}
    \shortintertext{and}
    \text{Mass equation: } \nabla \cdot \vec{u} =& 0
    \label{eq:divergence}
    \shortintertext{where}
    \rho =& \text{ fluid density,} \nonumber \\
    p =& \text{ pressure,} \nonumber \\
    \mu =& \text{ dynamic viscosity,} \nonumber \\
    \alpha =& \text{ coefficient of thermal expansion,} \nonumber \\
    \vec{g} =& \text{ gravitational force vector,} \nonumber
    \\
    T =& \text{ fluid temperature,} \nonumber \\
    T_{\text{ref}} =& \text{ reference temperature at which the nominal}
    \nonumber \\
    &\text{ density is provided.} \nonumber
    \nonumber
\end{align}

@@

\begin{align}
    \rho c_{p} \frac{\partial T}{\partial t} =& - \rho c_p \vec{u}
    \cdot \nabla T + \nabla \cdot \left(k \nabla T \right) + Q_f - Q_s
    \label{eq:temp}
    \shortintertext{and}
    Q_f =& \sum^G_{g=1} \epsilon_g \Sigma_g^f \phi_g \label{eq:heat-source}
    \shortintertext{where}
    c_p =& \text{ specific heat capacity of molten salt,} \nonumber \\
    k =& \text{ effective thermal conductivity of molten salt,} \nonumber \\
    Q_f =& \text{ fission heat source,} \nonumber \\
    \epsilon_g =& \text{ average fission energy released by neutrons in group
    $g$,} \nonumber \\
    Q_s =& \text{ heat sink/removal.} \nonumber
\end{align}

@@

Moltres also accounts for the decay of
\glspl{DNP} outside the active core region by simulating its flow in a
separate 1-D pipe geometry. This external loop pipe calculation is tightly
coupled to the active core simulation through Picard iterations in MOOSE's
MultiApp \cite{gaston_physics-based_2015} functionality and inlet/outlet boundary values.
The external loop region is assumed to be subcritical to minimize neutron
irradiation upon heat exchangers, pumps, and other equipment. Therefore, the
only significant neutronic-related phenomena are the drift and decay of
\glspl{DNP}. The governing equation for the \glspl{DNP} is:
%
\begin{align}
    \frac{\partial C_i}{\partial t} =& - \lambda_i C_i - u
    \frac{\partial C_i}{\partial x}.
    \label{eq:dnploop}
\end{align}
%
Equation \ref{eq:dnploop} is derived from equation \ref{eq:precursor} by
removing the fission \gls{DNP} source and diffusion terms and reducing the
dimensionality from 3-D to 1-D.
